General constraints on magic distillation protocols from statistical mechanics

### 2021-04-18 magic_paper.tex
Blending of new perspective. Bits need checking for logical/detail gaps.

We now turn to the application of majorization to describing magic in different fragments. The simplest way to express the conditions for interconversion is through Lorenz curves, which we now describe.

@@

\ddd{[I'll tidy this up more tomorrow and extend notation to full relative majorization.]}
However, the eagle-eyed reader will point out that relative majorization is usually considered for probability distributions, and so we verify
 that the same conditions apply to quasi-probability distributions. This can be done from first principles, but a simple way is to reduce to the problem of genuine distributions by `masking' the negativity in $\w$ using the reference distribution $\r$.

More precisely, if the components of $\r$ are never zero, then there always exists an $\epsilon >0$ such that $\w_\epsilon := \epsilon \w + (1-\epsilon) \r$ is a genuine probability distribution. We can see that $(\w , \r) \succ (\w', \r')$ if and only if $(\w_\epsilon , \r) \succ (\w_\epsilon', \r')$, where we choose $\epsilon$ so that both $\w_\epsilon$ and $\w_\epsilon'$ are probability distributions. This equivalence holds because there exists a stochastic map $A$ such that $A \w = \w'$ and $A \r = \r'$ if and only if $A[\epsilon \w + (1-\epsilon) \r] = \epsilon \w' + (1-\epsilon) \r'$ and $ A\r = \r'$. Now it is shown in \ddd{[cite ``The mixing distance'', E Ruch, R Schranner, TH Seligman]} relative majorization between pairs of distributions occurs if and only if the associated Lorenz curves never cross. Therefore we need only establish that this carries over to Lorenz curves for quasi-distributions.

In the appendices we show that 
\begin{equation}
L_{a \w + b \r | \r} (x) = a L_{\w|\r} (x) + bL _{\r|\r}(x),
\end{equation}
for all $a >0$ and any $b \in \mathbb{R}$. We now have that
\begin{equation}
(\w, \r) \succ (\w', \r') \mbox{ if and only if } (\w_\epsilon , \r) \succ (\w_\epsilon', \r'),
\end{equation}
and also $(\w_\epsilon , \r) \succ (\w_\epsilon', \r')$ if and only if $L_{\w_\epsilon |\r} (x) \ge L_{\w'_\epsilon |\r'} (x)$ for all $x$. However, using the above linearity property of the Lorenz curves we have that
\begin{align}
L_{\w_\epsilon |\r} (x) &\ge L_{\w'_\epsilon |\r'} (x) \nonumber \\ 
&\mbox{ if and only if }& \nonumber \\
\epsilon L_{\w |\r} (x) + (1-\epsilon) L_{\r|\r}(x) &\ge \epsilon L_{\w' |\r'} (x) + (1-\epsilon) L_{\r'|\r'}(x).
\end{align}
However $L_{\r|\r}(x)$ is simply the function $f(x)=x$ for any $\r$ and so the $(1-\epsilon)$ terms cancel on both sides and we get the required relative majorization conditions for the quasi-distributions in terms of their Lorenz curves.

@@

\section{Magic bounds in arbitrary stabiliser fragments and energetic costs of distillation}
\label{sec:stab}

We now generalise the approach taken in the previous section and consider bounds on magic distillation for an arbitrary stabiliser fragment $\R_\sigma$, where $\sigma$ is any quantum state $\sigma \in \stab$.
In other words, we consider those bounds on distillation that apply when the free operations have $\sigma$ as a fixed point.

These bounds can be interpreted in two different ways: 
\begin{enumerate}
\item As a family of upper bounds parameterised by an abstract stabiliser state $\sigma$.
\item As a constraint arising from the actual hardware limitations, noise bias or ambient thermal environment and reflected by the invariant state $\sigma$ for the system.
\end{enumerate}
In what follows we shall be agnostic on the interpretation. Without loss of generality, we can always write a full-rank stabiliser state $\sigma$ in the form $\sigma =  \frac{1}{\Z_\beta} e^{-\beta H}$ for some $\beta \geq 0$ and effective Hamiltonian $H$. Our results will also depend on the free energy $F(\beta)$ of the state $\sigma$, expressed as
\begin{equation}
	F(\beta) \coloneqq \tr[H \sigma] - \beta^{-1}S(\sigma) = -\beta^{-1}\log{\Z_\beta}.
\end{equation}
Here $S(\sigma)$ is the von Neumann entropy $S(\sigma) \coloneqq -\tr[\sigma\log{\sigma}]$. Note that under interpretation (1) above, this is an abstract entropic measure of the stabilizer state, while, under interpretation (2), it corresponds to the actual thermodynamic free energy of the equilibrium state.

We focus on the distillation process
\begin{equation}\label{eq:stdist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \sigma^{\otimes (n-n')},
\end{equation}
where, as before, tensoring in copies of $\sigma$ does not affect the Lorenz curves of the magic states in the fragment.

@@

\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.4]{figs/rate_scatter.pdf}
    \caption{\textbf{Magic distillation bounds within arbitrary stabiliser fragments.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ where $\beta$ is an effective inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ at every $\beta \leq \beta_\star$. The unital fragment corresponds to the $\beta =0 $ line.
    }
    \label{fig:rate_contour}
\end{figure}

\begin{theorem}\label{thm:stab_bounds}
	Given a theory of magic $\R = (\F, \O)$, let $\sigma$ be any qutrit stabiliser state, and write the state as $\sigma = \frac{1}{\Z} e^{-\beta H}$, where $H$ has eigenvalues $E_0, E_1, E_2$, and $\beta$ is an effective inverse temperature for the state. 
We define $\beta_\star = 1/kT_\star$ through the relation
\begin{equation}
	E_{\rm max} - E_s \eqqcolon kT_\star \ln 2,
\end{equation}
where $E_{\rm max} = \max\{E_0, E_1, E_2\}$ and $E_s$ is the eigenvalue of the eigenstate that overlaps the negative component of $\rho_S(\epsilon)$ in the Wigner representation. For $\beta \leq \beta_\star$, we define a threshold noise level
\begin{equation}
	\epsilon_{\star}(\beta) \coloneqq 3 - \dfrac{18}{8-e^{(E_{\rm max} - E_s)\beta}}.
\end{equation}
Then, any distillation rate $R(\epsilon, \beta)$ in the $\sigma$--fragment of a magic theory is bounded as follows:\\
If $\beta \leq \beta_{\star}$ and $\epsilon  \leq \epsilon_{\star}$,
\begin{equation}
	R(\epsilon,\beta) \leq 1 + \frac{\ln{\left( 1 - \frac{4}{3}\epsilon \right)}}{\beta (E_s - F(\beta))}.
\end{equation}
If $\beta \leq \beta_{\star}$ and $\epsilon_{\star} < \epsilon$, then
\begin{equation}
	R(\epsilon, \beta) \le 1 + \frac{\ln{\left(1-\frac{1}{3}\epsilon \right)} - (E_{\rm max} - E_s)(\beta_{\star} - \beta)}{\beta (E_s - F(\beta))}.
\end{equation}
Otherwise if $\beta > \beta_{\star}$,
\begin{equation}
	R(\epsilon, \beta) \leq  1+ \frac{\ln{\left(1-\frac{1}{3}\epsilon \right)}}{-\ln{2} + \beta (E_{\rm max} - F(\beta))}.
\end{equation}
In the above expressions $F(\beta)$ is the effective free energy of the stabilizer state.
\end{theorem}
A few comments can be given on this result. 
Firstly, the specific numerical factors in $\epsilon_\star$ are a result of our choice of magic state. 
Secondly, these bounds are derived based on analysis of only a part of the Lorenz curves and can be improved via a finer analysis. 
This is apparent by simple numerical calculations on the entirety of the curves.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure. 
Finally, it is notable that we obtain a Landauer-like condition with a characteristic temperature $kT_\star \ln 2 = E_{\rm max} - E_s$, where $kT_\star = \beta_\star^{-1}$. 
It is unclear whether this points to a generic feature that can be directly related to fundamental thermodynamic relations, such as the erasure cost of a single bit being $kT \ln 2$. 
Given that we work with qutrits this seems surprising, but does deserve further study.

@@

Using that $F(\beta) = -\beta^{-1} \log \Z$, the first elbow coordinates in the two cases are now given by
\begin{equation}\label{eq:first_elb_coords}
	(x_0, L_0) =
	\begin{cases}
		\left(\frac{1}{3^n} e^{-n\beta (E_s - F(\beta))}, v^n \right), &\rm{(C1)}	\vspace{10pt}\\
		\left( e^{-n\beta (E_{\rm{max}}-F(\beta))}, (3u)^n \right). &\rm{(C2)} 
	\end{cases}
\end{equation}
To obtain the distillation bounds from majorization we can now compute the above terms for both the input and output states. Depending on the noise parameters $\epsilon, \epsilon'$ for the input and output states the Lorenz curve data may fall into either case $\rm{(C1)}$ or case $\rm{(C2)}$.
Specifically, we have three scenarios of relevance to magic distillation:
\begin{enumerate}
	\item $\rm{(C1)} \rightarrow \rm{(C1)}$ if $E_{\rm{max}} = E_s$ or $E_{\rm{max}} > E_s$, $\beta < \beta_{\star}$ and $\epsilon' < \epsilon  \leq \epsilon_{\star}$.
	\item $\rm{(C2)} \rightarrow \rm{(C1)}$ if $E_{\rm{max}} > E_s$, $\beta < \beta_{\star}$ and $\epsilon' \leq \epsilon_{\star} < \epsilon$.
	\item $\rm{(C2)} \rightarrow \rm{(C2)}$ if $E_{\rm{max}} > E_s$, $\beta < \beta_{\star}$ and $\epsilon_{\star} \leq \epsilon' < \epsilon$ or $E_{\rm{max}} > E_s$, $\beta \geq \beta_{\star}$.
\end{enumerate}
The scenario $\rm{(C1)} \rightarrow \rm{(C2)}$ is not considered as it would correspond to $\epsilon \leq \epsilon_\star < \epsilon'$, which would imply that the output magic state is noisier than the input state.

In each of the three scenarios, we simply use the first elbow constraint, as described in~\cref{app:elb_constraints}, which leads to the following distillation bounds,
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq
	\begin{cases}
		\dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (E_s - F(\beta))}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E_s - F(\beta))},\ &\rm{(C1)} \rightarrow \rm{(C1)}, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F(\beta))}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E_s - F(\beta))},\ &\rm{(C2)} \rightarrow \rm{(C1)}, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F(\beta))}{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon' \big)} + \beta (E_{\rm{max}} - F(\beta))},\ &\rm{(C2)} \rightarrow \rm{(C2)}.
	\end{cases}
\end{equation}
Substituting $\epsilon' = 0$ into these expressions leads to the stated distillation bounds.

@@

We consider an arbitrary change in the Hamiltonian driving the system, $H\rightarrow H'$ such that $\sigma' = e^{-\beta H'}/\Z'$, so the energetic parameters contributing to the Lorenz curve of the output state $\rho'$ relative to $\sigma'$ will change. 
Specifically, the first elbow coordinates of the output Lorenz curve can now be expressed as
\begin{equation}\label{eq:transformed_first_elb_coords}
	(x'_0, L'_0) =
	\begin{cases}
		\left(\frac{1}{3^{n'}} e^{-n\beta (E'_s - F'(\beta))}, v(\epsilon')^{n'} \right), &\hspace{-7pt} \rm{(C1')}	\vspace{10pt}\\
		\left( e^{-n'\beta (E'_{\rm{max}}\hspace{-2.5pt}-F'(\beta))}, (3u(\epsilon'))^{n'} \right), &\hspace{-7pt} \rm{(C2')}
	\end{cases}
\end{equation}
where now Case $\rm{(C1')}$ applies when $\epsilon' \leq \epsilon'_\star(\beta'_\star)$ (which encompasses the case $E'_s = E'_{\rm{max}}$) and Case $\rm{(C2')}$ applies when $\epsilon' > \epsilon'_\star(\beta'_\star)$ (which encompasses the case $\beta \geq \beta'_\star$).
It is therefore now possible to get the scenario $\rm{(C1)} \rightarrow \rm{(C2')}$ if the two Hamiltonians $H,H'$ are sufficiently different.

Using the first elbow constraint again leads to the following distillation bounds,
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq
	\begin{cases}
		\dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (E_s - F(\beta))}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E'_s - F'(\beta))},\ &\rm{(C1)} \rightarrow \rm{(C1')}, \vspace{10pt}\\
		\dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (E_s - F(\beta))}{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon' \big)} + \beta (E'_{\rm{max}} - F'(\beta))},\ &\rm{(C1)} \rightarrow \rm{(C2')}, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F(\beta))}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E'_s - F'(\beta))},\ &\rm{(C2)} \rightarrow \rm{(C1')}, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F(\beta))}{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon' \big)} + \beta (E'_{\rm{max}} - F'(\beta))},\ &\rm{(C2)} \rightarrow \rm{(C2')}.
	\end{cases}
\end{equation}

>>>>

\ddd{[Notation needs smoothing to accommodate new points of emphasis]}
We now turn to the application of majorization to describing magic in different fragments. The simplest way to express the conditions for interconversion is through Lorenz curves, which we now describe.

@@

However, relative majorization is usually considered for probability distributions, and so we must verify that the same Lorenz curve conditions apply to quasi-probability distributions. This can be done from first principles, but a simpler way is see that this holds is to reduce to the problem of genuine distributions by first `masking' the negativity in the quasi-probability distribution $\w$ using the reference distribution $\r$ and then applying the conditions for relative majorization.

More precisely, if the components of $\r$ are never zero, then there always exists an $\epsilon >0$ such that $\w_\epsilon := \epsilon \w + (1-\epsilon) \r$ is a genuine probability distribution. We can see that $(\w , \r) \succ (\w', \r')$ if and only if $(\w_\epsilon , \r) \succ (\w_\epsilon', \r')$, where we choose $\epsilon$ so that both $\w_\epsilon$ and $\w_\epsilon'$ are probability distributions. This equivalence holds because there exists a stochastic map $A$ such that $A \w = \w'$ and $A \r = \r'$ if and only if 
\begin{equation}
A[\epsilon \w + (1-\epsilon) \r] = \epsilon \w' + (1-\epsilon) \r'\mbox{ and } A\r = \r'.
\end{equation}
Now it is shown in \ddd{[cite ``The mixing distance'', E Ruch, R Schranner, TH Seligman]} relative majorization between pairs of distributions occurs if and only if the associated Lorenz curves never cross. Therefore we can express this condition in terms of the Lorenz curves of the quasi-probability distributions $\w$ and $\w'$ relative to $\r$ and $\r'$ respectively.

As shown in Appendix \ref{lemma:Lorenz_linearity}, the Lorenz curve for any quasi-probability distribution obeys the following linearity condition
\begin{equation}
L_{a \w + b \r | \r} (x) = a L_{\w|\r} (x) + bx,
\end{equation}
for all $a >0$ and any $b \in \mathbb{R}$. We now have that
\begin{equation}
(\w, \r) \succ (\w', \r') \mbox{ if and only if } (\w_\epsilon , \r) \succ (\w_\epsilon', \r'),
\end{equation}
and also $(\w_\epsilon , \r) \succ (\w_\epsilon', \r')$ if and only if $L_{\w_\epsilon |\r} (x) \ge L_{\w'_\epsilon |\r'} (x)$ for all $x$. However, using the above linearity property of the Lorenz curves we have that
\begin{align}
L_{\w_\epsilon |\r} (x) &\ge L_{\w'_\epsilon |\r'} (x) \nonumber \\ 
&\mbox{ if and only if }& \nonumber \\
\epsilon L_{\w |\r} (x) + (1-\epsilon) x &\ge \epsilon L_{\w' |\r'} (x) + (1-\epsilon) x.
\end{align}
The $(1-\epsilon)$ terms cancel on both sides and we get the required relative majorization conditions for the quasi-distributions in terms of their Lorenz curves.

@@

\section{Free energy dependent bounds for magic state distillation protocols}
\label{sec:stab}
\ddd{[Various smoothings needed to blend with new emphasis]}
%
%We now generalise the approach taken in the previous section and consider bounds on magic distillation for an arbitrary stabiliser fragment $\R_\sigma$, where $\sigma$ is any quantum state $\sigma \in \stab$.
%In other words, we consider those bounds on distillation that apply when the free operations have $\sigma$ as a fixed point.
%
%These bounds can be interpreted in two different ways: 
%\begin{enumerate}
%\item As a family of upper bounds parameterised by an abstract stabiliser state $\sigma$.
%\item As a constraint arising from the actual hardware limitations, noise bias or ambient thermal environment and reflected by the invariant state $\sigma$ for the system.
%\end{enumerate}
%In what follows we shall be agnostic on the interpretation.

Without loss of generality, we can always write a full-rank stabiliser state $\sigma$ in the form $\sigma =  \frac{1}{\Z_\beta} e^{-\beta H}$ for some $\beta \geq 0$ and effective Hamiltonian $H$. Our results will also depend on the free energy $F$ of the state $\sigma$, expressed as
\begin{equation}
	F \coloneqq \tr[H \sigma] - \beta^{-1}S(\sigma) = -\beta^{-1}\log{\Z_\beta}.
\end{equation}
Here $S(\sigma)$ is the von Neumann entropy $S(\sigma) \coloneqq -\tr[\sigma\log{\sigma}]$. Note that under interpretation (1) above, this is an abstract entropic measure of the stabilizer state, while, under interpretation (2), it corresponds to the actual thermodynamic free energy of the equilibrium state.
%
%We focus on the distillation process
%\begin{equation}\label{eq:stdist}
%	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \sigma^{\otimes (n-n')},
%\end{equation}
%where, as before, tensoring in copies of $\sigma$ does not affect the Lorenz curves of the magic states in the fragment.

@@

\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.4]{figs/rate_scatter.pdf}
    \caption{\textbf{Magic distillation bounds within arbitrary stabiliser fragments.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ where $\beta$ is an effective inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ at every $\beta \leq \beta_\star$. The unital fragment corresponds to the $\beta =0 $ line.
    }
    \label{fig:rate_contour}
\end{figure}
%
%\begin{theorem}\label{thm:stab_bounds}
%	Given a theory of magic $\R = (\F, \O)$, let $\sigma$ be any qutrit stabiliser state, and write the state as $\sigma = \frac{1}{\Z} e^{-\beta H}$, where $H$ has eigenvalues $E_0, E_1, E_2$, and $\beta$ is an effective inverse temperature for the state. 
%We define $\beta_\star = 1/kT_\star$ through the relation
%\begin{equation}
%	E_{\rm max} - E_s \eqqcolon kT_\star \ln 2,
%\end{equation}
%where $E_{\rm max} = \max\{E_0, E_1, E_2\}$ and $E_s$ is the eigenvalue of the eigenstate that overlaps the negative component of $\rho_S(\epsilon)$ in the Wigner representation. For $\beta \leq \beta_\star$, we define a threshold noise level
%\begin{equation}
%	\epsilon_{\star}(\beta) \coloneqq 3 - \dfrac{18}{8-e^{(E_{\rm max} - E_s)\beta}}.
%\end{equation}
%Then, any distillation rate $R(\epsilon, \beta)$ in the $\sigma$--fragment of a magic theory is bounded as follows:\\
%If $\beta \leq \beta_{\star}$ and $\epsilon  \leq \epsilon_{\star}$,
%\begin{equation}
%	R(\epsilon,\beta) \leq 1 + \frac{\ln{\left( 1 - \frac{4}{3}\epsilon \right)}}{\beta (E_s - F)}.
%\end{equation}
%If $\beta \leq \beta_{\star}$ and $\epsilon_{\star} < \epsilon$, then
%\begin{equation}
%	R(\epsilon, \beta) \le 1 + \frac{\ln{\left(1-\frac{1}{3}\epsilon \right)} - (E_{\rm max} - E_s)(\beta_{\star} - \beta)}{\beta (E_s - F)}.
%\end{equation}
%Otherwise if $\beta > \beta_{\star}$,
%\begin{equation}
%	R(\epsilon, \beta) \leq  1+ \frac{\ln{\left(1-\frac{1}{3}\epsilon \right)}}{-\ln{2} + \beta (E_{\rm max} - F)}.
%\end{equation}
%In the above expressions $F$ is the effective free energy of the stabilizer state.
%\end{theorem}

\ddd{[Put assumptions here. Needs more polish+checking.]} We now consider a magic distillation protocol on multiple identical qutrits sending 
\begin{equation}
\rho_S(\epsilon)^{\otimes n} \longrightarrow \rho_S(\epsilon')^{\otimes Rn}
\end{equation}•
with $n \rightarrow \infty$. Let each qutrit have a Hamiltonian $H$ with energies $E_0, E_1, E_2$. Let $T =(k\beta)^{-1}$ be any characteristic temperature for the physical system.

We can now provide distillation bounds that depend on energetic and entropic aspects of the protocol. However, this is done relative to a reference temperature that can be chosen freely. The measures associated with this protocol are then determined by how it would disturb this reference equilibrium state $\sigma$ to some different state $\sigma'$. More precisely, the reference equilibrium state of the $n$ qutrits is given by
\begin{equation}
\sigma^{\otimes n} = \left ( \frac{e^{-\beta H}}{\Z} \right )  ^{\otimes n}.
\end{equation}
Since we are concerned with magic distillation we may assume that the reference state $\sigma$ is not a magic state, and so has non-negative Wigner distribution. Instead we shall assume that it is a full-rank stabilizer state for the qutrit. A given magic protocol on the $n$ qutrits will correspond to a quantum channel $\E$ acting on $\rho_S(\epsilon)^{\otimes n}$. Our analysis associates to this protocol energy/entropy measures by considering how $\E$ transforms this reference equilibrium state. For simplicity we assume that the action of $\E$ on $\sigma^{\otimes n}$ is also such that correlations can be neglected in the $n\rightarrow \infty$ limit, and so $\E(\sigma^{\otimes n}) = (\sigma')^{\otimes n}$ for some state $\sigma'$. This assumption simplifies the application of majorization, however we note that work has also been done on the case of where correlations build up and one only looks at the marginal state on an individual subsystem \ddd{[need to cite here]} and so a more detailed analysis can also be performed.

\begin{theorem}\label{thm:free-energy}
	Consider a magic distillation protocol on $n \gg 1$ identical qutrits under the assumptions just stated. Let $T =(k\beta)^{-1}$ be any characteristic temperature for the physical system.

\noindent Define $\beta_\star = 1/kT_\star$ through the relation
\begin{equation}
	E_{\rm max} - E_s \eqqcolon kT_\star \ln 2,
\end{equation}
where $E_{\rm max} = \max\{E_0, E_1, E_2\}$ and $E_s$ is the eigenvalue of the eigenstate that overlaps the negative component of $\rho_S(\epsilon)$ in the Wigner representation. For $\beta \leq \beta_\star$, we define a threshold noise level
\begin{equation}
	\epsilon_{\star}(\beta) \coloneqq 3 - \dfrac{18}{8-e^{(E_{\rm max} - E_s)\beta}}.
\end{equation}
Then the distillation rate $R$ of the protocol is bounded as follows:
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq
	\begin{cases}
		\dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (E_s - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E'_s - F')},\ &\rm{(C1)} \rightarrow \rm{(C1')}, \vspace{10pt}\\
		\dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (E_s - F)}{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon' \big)} + \beta (E'_{\rm{max}} - F')},\ &\rm{(C1)} \rightarrow \rm{(C2')}, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E'_s - F')},\ &\rm{(C2)} \rightarrow \rm{(C1')}, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F)}{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon' \big)} + \beta (E'_{\rm{max}} - F')},\ &\rm{(C2)} \rightarrow \rm{(C2')}.
	\end{cases}
\end{equation}
In the above expressions $F$ is the free energy of the state $\sigma:=e^{-\beta H}/\Z$, $F'$ is the free energy of the single qutrit state $\sigma' =e^{-\beta H'}/\Z'$ defined via the action of the protocol on the reference equilibrium state. \ddd{[Need strategy to incorporate conditions in simple, clear way.]}
\end{theorem}
 The state $\sigma'$ is generally a non-equilibrium state for the system, however one can still associate a free energy value to the output by considering subsequently changing the Hamiltonian $H \rightarrow H'$ such that equilibrium is restored at the reference temperature $T$. For example, suppose that $H=H'$ then this implies the protocol would leave the equilibrium state unchanged and so corresponds to a Gibbs-preserving map \ddd{[cite Philippe]}. In the more general case $\Delta F := F' - F \ne 0$ and we have a free energy difference between the initial and final times, with respect to the relevant Hamiltonians. This is similar to the free energy differences that occur in fluctuation relations, where the final state of the system can be arbitrarily far from equilibrium, yet a free energy term for the final system still arises.

Firstly, the specific numerical factors in $\epsilon_\star$ are a result of our choice of magic state. 
Secondly, these bounds are derived based on analysis of only a part of the Lorenz curves and can be improved via a finer analysis. 
This is apparent by simple numerical calculations on the entirety of the curves.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure.

@@

Using that $F = -\beta^{-1} \log \Z$, the first elbow coordinates in the two cases are now given by
\begin{equation}\label{eq:first_elb_coords}
	(x_0, L_0) =
	\begin{cases}
		\left(\frac{1}{3^n} e^{-n\beta (E_s - F)}, v^n \right), &\rm{(C1)}	\vspace{10pt}\\
		\left( e^{-n\beta (E_{\rm{max}}-F)}, (3u)^n \right). &\rm{(C2)} 
	\end{cases}
\end{equation}
To obtain the distillation bounds from majorization we can now compute the above terms for both the input and output states. Depending on the noise parameters $\epsilon, \epsilon'$ for the input and output states the Lorenz curve data may fall into either case $\rm{(C1)}$ or case $\rm{(C2)}$.
Specifically, we have three scenarios of relevance to magic distillation:
\begin{enumerate}
	\item $\rm{(C1)} \rightarrow \rm{(C1)}$ if $E_{\rm{max}} = E_s$ or $E_{\rm{max}} > E_s$, $\beta < \beta_{\star}$ and $\epsilon' < \epsilon  \leq \epsilon_{\star}$.
	\item $\rm{(C2)} \rightarrow \rm{(C1)}$ if $E_{\rm{max}} > E_s$, $\beta < \beta_{\star}$ and $\epsilon' \leq \epsilon_{\star} < \epsilon$.
	\item $\rm{(C2)} \rightarrow \rm{(C2)}$ if $E_{\rm{max}} > E_s$, $\beta < \beta_{\star}$ and $\epsilon_{\star} \leq \epsilon' < \epsilon$ or $E_{\rm{max}} > E_s$, $\beta \geq \beta_{\star}$.
\end{enumerate}
The scenario $\rm{(C1)} \rightarrow \rm{(C2)}$ is not considered as it would correspond to $\epsilon \leq \epsilon_\star < \epsilon'$, which would imply that the output magic state is noisier than the input state.

In each of the three scenarios, we simply use the first elbow constraint, as described in~\cref{app:elb_constraints}, which leads to the following distillation bounds,
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq
	\begin{cases}
		\dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (E_s - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E_s - F)},\ &\rm{(C1)} \rightarrow \rm{(C1)}, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E_s - F)},\ &\rm{(C2)} \rightarrow \rm{(C1)}, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F)}{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon' \big)} + \beta (E_{\rm{max}} - F)},\ &\rm{(C2)} \rightarrow \rm{(C2)}.
	\end{cases}
\end{equation}
Substituting $\epsilon' = 0$ into these expressions leads to the stated distillation bounds.

@@

We consider an arbitrary change in the Hamiltonian driving the system, $H\rightarrow H'$ such that $\sigma' = e^{-\beta H'}/\Z'$, so the energetic parameters contributing to the Lorenz curve of the output state $\rho'$ relative to $\sigma'$ will change. 
Specifically, the first elbow coordinates of the output Lorenz curve can now be expressed as
\begin{equation}\label{eq:transformed_first_elb_coords}
	(x'_0, L'_0) =
	\begin{cases}
		\left(\frac{1}{3^{n'}} e^{-n\beta (E'_s - F')}, v(\epsilon')^{n'} \right), &\hspace{-7pt} \rm{(C1')}	\vspace{10pt}\\
		\left( e^{-n'\beta (E'_{\rm{max}}\hspace{-2.5pt}-F')}, (3u(\epsilon'))^{n'} \right), &\hspace{-7pt} \rm{(C2')}
	\end{cases}
\end{equation}
where now Case $\rm{(C1')}$ applies when $\epsilon' \leq \epsilon'_\star(\beta'_\star)$ (which encompasses the case $E'_s = E'_{\rm{max}}$) and Case $\rm{(C2')}$ applies when $\epsilon' > \epsilon'_\star(\beta'_\star)$ (which encompasses the case $\beta \geq \beta'_\star$).
It is therefore now possible to get the scenario $\rm{(C1)} \rightarrow \rm{(C2')}$ if the two Hamiltonians $H,H'$ are sufficiently different.

Using the first elbow constraint again leads to the following distillation bounds,
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq
	\begin{cases}
		\dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (E_s - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E'_s - F')},\ &\rm{(C1)} \rightarrow \rm{(C1')}, \vspace{10pt}\\
		\dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (E_s - F)}{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon' \big)} + \beta (E'_{\rm{max}} - F')},\ &\rm{(C1)} \rightarrow \rm{(C2')}, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (E'_s - F')},\ &\rm{(C2)} \rightarrow \rm{(C1')}, \vspace{10pt}\\
		\dfrac{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon \big)} + \beta (E_{\rm{max}} - F)}{\ln{\big( \frac{1}{2}-\frac{1}{6}\epsilon' \big)} + \beta (E'_{\rm{max}} - F')},\ &\rm{(C2)} \rightarrow \rm{(C2')}.
	\end{cases}
\end{equation}
